Parametric controls on vegetation responses to biogeochemical forcing in the CLM5

### 2019-06-24 referee_responses.tex

So, my suggestion here would be to show the impact of each of the 3(4) new assumptions in relation to how nutrients are treated with CLM5 and then to explore how sensitive those results were to the underlying params. Here, my expectation was that in some cases the results would be insensitive, but in others, it is clear that the notion of a single PFT value does not make sense. As I understand it, the authors varied the gs slope term based on the param range in my 2015 paper. This gives them some erroneous results, but that wouldn't be unexpected as using the C3 grass param (a "spendy" plant) in an arid environment wouldn't make biophysical sense?

\textsf{AR: We agree with this idea wrt. the range of variation across PFTs perhaps being excessively wide. We have now modified the stomatal slope range to use the standard error reported in Table 2 of de Kauwe et al. 2015}}

In my opinion the editor shouldn't impose such a change on the authors as it is potentially a lot of work and not what they wished their paper to be about. It is just my suggestion. Equally, I would have loved to have seen those N hypotheses tested at the Duke/Oak Ridge FACE experiments, perhaps that is for another paper :P.

\textsf{AR: We actually ran the model for those sites, but got very low LAI predictions, and therefore very sensitive responses to fertilization. We thought in the end that fixing up the model to perform better at those particular sites was less interesting than the overall controls over the response to CO$_{2}$ and N forcing.}

### 2019-06-26 main_revision.tex

Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their fixation of nitrogen. Model initial state, stoichiometric flexibility of plant tissues, and nitrogen uptake costs all have notable impacts, but in general the output space is complex, and varies between sites, output variables, and fertilization scenarios.  These responses can typically be understood in the context of the model structure and the functionality of individual parameters. Identification of the primary model features driving ecosystem responses to CO$_{2}$  and N fertilization is a necessary step in the interpretation of global carbon cycle feedback simulations, and in the identification of priorities for fieldwork and process updates. Further studies focused on how alternative parameters impact biogeochemical function would provide useful compliments to the ongoing CMIP6 process.
\end{abstract}

\section{Introduction}
Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model inter-comparison, one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth. In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}).

The Community Land Model is the land surface representation of the Community Earth System Model (CESM, \cite{hurrell2013}). Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019} and plant hydraulics by \cite{kennedy2019}. Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.

The CLM5 includes numerous major updates of the nitrogen cycle representation relative to CLM4 and CLM4.5 (\cite{lawrence2011}). The introduction of several new dynamical features into the CLM5 has implications for responses to environmental forcing. \cite{wieder2019} report the global-scale carbon cycle implications of the full model update and note that, in the default CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.

@@

1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control model state.

2. There is a trade-off between model parameterizations that allow strong CO2 fertilization and those that allow strong N fertilization.

The analysis herein does not focus on the relative fit of the model to the data collected at the individual sites. The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics, a focus on that would preclude detailed analysis of the parameter space, as presented here. Assessment of the skill of the CLM5 model against a broad suite of globally-relevant data products and manipulation experiments is the focus of \cite{lawrence2018} and \cite{wieder2019}. Future studies will consider the impact of alternative parameterizations, informed by these analyses, on the global performance and skill of CLM5.

@@

@@

Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}). Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}), which appears to be a poor means of capturing its variation in time and space and responses to climatic change (\cite{walker2017}). There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).  LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for a given leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.  $V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5. FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun} and \cite{brzostek2014} and \cite{shi2016}). A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}). Previous versions of CLM model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}). The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}). Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake.   The switching of N uptake to cheaper pathways results in the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}), whereas CLM4 and CLM4.5 predicted N fixation as a linear function of net primary productivity (\cite{cleveland1999}; \cite{wieder2015}).   The FUN model is documented within the CLM technical note at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}.

@@

For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation. Following \cite{wieder2019}, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998. Similar to \cite{wieder2019}, we assessed the impact of changes after 15-19 years of fertilization in both cases.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.  Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization. These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in it's own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.  We focus here on biogeochemically important inputs. Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme will be discussed by \cite{kennedy2019}.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around the default state, which (compared to a global parameter sensitivity analysis) allows more straightforward visualization and interpretation of the results.  At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below. We recognize that a global parameter sensitivity analysis would also be of interest, particularly to identify interacting effects, and indeed have conducted many such analyses. We concluded that a OAAT analysis provided clearer insight into the functionality of the parameters with respect to biogeochemical forcing, and further that the more complex visualization and analysis of a global sensitivity study, across the entire domain of results presented here, would reduce the opportunity for understanding the individual parameter responses in depth, as is the goal of our study.

@@

The parameters included in our sensitivity analysis are listed in Table \ref{table_ranges}. Their impact on model functionality is described below. As before, further details can be found in the appendix to \cite{lawrence2018}. The shorthand version of the parameter name is given in italics.

\textbf{Specific Leaf Area}, \emph{`SLA'}, is a parameter common to almost all land surface models, which determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light. In CLM5 high SLA leaves by definition also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.

\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass. In the CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow (\cite{kennedy2019}) and the cost of N acquisition in the FUN model, respectively. The impact of this parameter is thus of greater interest than its functionality as a simple 'tax' on plant growth in previous versions of the CLM. Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.  The carbon allocation  in CLM5 uses a fixed-fraction approach to allocation in general, with constant ratios between, leaf, fine roots, stem and coarse root allocation. We note here that the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance, an feature which must be considered when analyzing simulations under high productivity conditions. Newer codes within the CLM framework (e.g. the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}) account for this issue by removing leaves in negative carbon balance.

 \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the allocation of biomass to stem and woody tissues, which in CLM5 is again a constant fraction of leaf allocation. Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass. Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter, a feature which will be resolved in future model versions that allow plants to compete for light resources (\cite{fisher2018}).

\textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils.  FUN introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake. Therefore, uncertainties associated with the cost functions, in particular, are large. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The six \emph{'n\_costs'} parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.

@@

\section{Results}
Our analysis used four sites, with thirteen parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations. Including the default simulations, the total comprises 672 different run configurations. We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

@@

\subsubsection{Water and N limitation across sites}
The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}). Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential (\cite{kennedy2019}). $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$. Given that leaf water potential and  $B_{tran}$ vary diurnally, we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure). For N limitation, the fraction of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake is a direct indicator of how N stress affects the carbon budget.

@@

\subsection{Parametric control over system state}
At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected since it is used directly in the calculation of both of these quantities. For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity). The impact of SLA on NPP can, however, be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI. At Niwot Ridge, the parameter range straddles an optimal SLA (or local maximum at the second-largest parameter values) for NPP, again reflecting the assimilation/respiration cost trade-off.

@@

Changing \emph{leafcn} -the target leaf C:N ratio- over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).  We do not find much evidence of saturation in the GPP vs. \emph{leafcn} relationships, in that GPP response does not level off at high leaf N content, in line with findings in \cite{lawrence2018} that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result. At Niwot, an optimum \emph{leafcn} ratio (or a local maximum in the parameteric response) was apparent. For NPP this was around the default value, but for LAI it was 0.5 standard deviations lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

@@

Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux. Water limitation is also modified as a result of altered transpiration. At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem. In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained. At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance. Note that \cite{kennedy2019} hydrodynamic functioning of the model in greater depth and that  studies currently underway will discuss surface energy and plant.

@@

\subsection{Differences across sites}
The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.

@@

\subsection{Impact of parameters}
Across the four sites, parametric impacts on fertilization responses were very varied, when compared to the relatively consistent impact of parameters on the initial system state. A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses across sites, and the range of their impacts varied over an order of magnitude. Thus, the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change. Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.

>>>>

 The impact of fertilization is dependant both on parametric variation explored here varies with the balance of limiting factors (water, temperature, nutrients, light) in the pre-fertilization model state.  The model parameters that impact the pre-fertilization state are in general not the same as those that control fertilization responses, meaning that calibration to present-day conditions will not necessarily constrain future projections. Where pre-fertilization state has low leaf area, fertilization induced increases in leaf production amplify the model response to the initial fertilization.  Model responses to CO$_{2}$ and N fertilization are strongly impacted by how much plant communities can increase their fixation of nitrogen.  Further studies focused on how alternative parameters impact biogeochemical function would provide useful compliments to the ongoing CMIP6 process.
\end{abstract}

\section{Introduction}
Simulating the cycling of nutrients, and how nutrient availability impacts ecosystem growth and function, has been repeatedly identified as a crucial element of Earth system models (\cite{piao2013}, \cite{gruber2008}, \cite{wang2009}). In the CMIP5 multi-model inter-comparison, one land model, the Community Land Model v4 (CLM4), included nitrogen (N) cycling and limitation of plant growth. In CLM4 projections, the inclusion of N dynamics significantly limits the ability of the terrestrial biosphere to respond to fertilization by increasing atmospheric carbon dioxide (\cite{friedlingstein2006}, \cite{thornton2007} \cite{friedlingstein2014}, \cite{arora2013}), leading to speculation that the addition of nitrogen limitation as a more general feature of models might reduce the capacity for vegetation to respond to CO2 fertilization and hence reduce the degree to which the terrestrial biosphere can mitigate increasing atmospheric CO$_{2}$.

The newly-released CLM version 5  (CLM5,\cite{lawrence2018}) includes numerous major updates relative to CLM4 and CLM4.5 (\cite{lawrence2011}). The introduction of several new dynamical features, including substantial modification of the vegetation nitrogen cycling, as well as the introduction of plant hydrodynamics, altered surface hydrology and dynamic crop representation, has implications for responses to environmental forcing. \cite{wieder2019} report the global carbon cycle implications of these changes and note that, in CLM5, carbon uptake responds more strongly to CO$_{2}$ and less strongly to N fertilization than its predecessors, with both fertilization responses being closer to experimental results than the previous versions of the model.

@@

1. The parameters that dominate responses to CO$_{2}$ and nitrogen fertilization are distinct from those that control the pre-fertilization model state.

2. There is a trade-off between model parameterizations that allow strong CO$_{2}$ fertilization and those that allow strong N fertilization.

The analysis herein does not focus on the relative fit of the model to the data collected at the individual sites. The fit at individual sites of the globally adjusted model typically cannot be seen as indicative or otherwise of its skill in the absence of specific parameterization using the local vegetation, soil and hydrological characteristics. A focus on model fit at individual sites would preclude detailed consideration of how parameter variation impacts fertilization responses, as presented here. Assessment of the skill of the CLM5 model against a broad suite of globally-relevant data products and manipulation experiments is the focus of \cite{lawrence2018} and \cite{wieder2019}. Future studies will consider the impact of alternative parameterizations, informed by these analyses, on the global performance and skill of CLM5.

@@

\subsection{CLM5}
 Version 5 of the CLM was released to the research community in 2018 as part of the CESM2.0 release, and open-source releases and development are housed at \url{https://github.com/ESCOMP/ctsm/}. An overview of model developments is presented within this special issue by \cite{lawrence2018}, a description of the global nitrogen cycling by \cite{wieder2019}/. The plant hydraulics model is documented by \cite{kennedy2019}. Further manuscripts will describe the crop, urban and land use models, sensitivity to forcing and calibration of fast-timescale processes.

@@

Firstly, the  `LUNA' module (Leaf Utilization of Nitrogen for Assimilation),  simulates distribution of N between different leaf assimilation processes (\cite{xu2012}, \cite{ali2016}). Most land surface models predict photosynthetic parameters directly from leaf nitrogen content (\cite{kattge2009}, \cite{bonan2012}). This method, however, appears to be a poor means of capturing variation in photosynthetic capacity in time and space and responses to climatic change (\cite{walker2017}). There is significant evidence that plant photosynthetic capacity can respond to environmental conditions, including CO$_{2}$ fertilization (\cite{ainsworth2007}), changing temperature (\cite{hikosaka2005}), irradiance (\cite{niinemets1998}) and soil moisture (\cite{keenan2009}).  LUNA predicts the optimal partitioning of N among the maximum rate of carboxylation $V_{c,max}$, the maximum rate of electron transport $J_{max}$, and other leaf N components for a given leaf nitrogen (per unit area) and environmental boundary conditions (CO$_{2}$, temperature, humidity, soil moisture, radiation and day length), assuming co-limitation of photosynthesis by both electron capture and carboxylation processes.  $V_{c,max}$ is thus primarily controlled by the N per unit leaf area $N_{area}$ and the prevailing meteorology. The capacity of the LUNA model to represent appropriate responses to temperature and CO$_{2}$ was demonstrated by \cite{xu2012} while the global calibration and the geographical predictions of the model were described by \cite{ali2016}.

Secondly, the `FUN' (Fixation and Uptake of Nitrogen) model was incorporated into CLM5. FUN simulates the dynamics and carbon economics of N acquisition from the environment, building on (\cite{fisher2010fun} and \cite{brzostek2014} and \cite{shi2016}). A major issue confronting N cycle models is how to manage 'excess' carbon under conditions where carbon assimilation is faster than that which can be supported by nutrient uptake (\cite{dekauwe2014}). Previous versions of CLM model reduced the gross photosynthetic flux down to the level at which growth could be supported by N uptake. This process occurred after the calculation of stomatal conductance (which is linked to assimilation rates via the Ball-Berry model) and therefore created inconsistencies between C assimilation and water cycling, especially under conditions of N limitation, as discussed extensively in the literature (\cite{medlyn2011}, \cite{bonan2012}, \cite{dekauwe2014}, \cite{walker2014}). The FUN model, in contrast, employs the principle that for the alternative sources of nitrogen acquisition by plants (active uptake from soil, retranslocation from senescent tissues, and symbiotic N fixation) there is a concurrent cost in terms of carbon (\cite{bloom1985}, \cite{jiang2017}, \cite{terrer2018}) which is determined by the abundance of nitrogen in the soils and plant tissues and the temperature-dependant cost of N fixation. Within FUN, plants spend excess carbon for nitrogen uptake preferentially using pathways which are `cheaper' to them. Where N is scarce in the environment, carbon that would otherwise have been used for plant growth can be deployed to acquire (pay for) more N. Thus, N limitation leads to both reduced growth and higher expenditure of C on N uptake.   The switching of N uptake to cheaper pathways results in the preferential use of symbiotic N fixation when soil mineral N concentrations are low (\cite{vitousek2002}), whereas CLM4 and CLM4.5 predicted N fixation as a linear function of net primary productivity (\cite{cleveland1999}; \cite{wieder2015}).   The FUN model is documented within the CLM technical note at \url{https://escomp.github.io/ctsm-docs/doc/build/html/tech_note/FUN/CLM50_Tech_Note_FUN.html}.

@@

For each site/parameter/value combination, we also ran one elevated CO$_{2}$ and one elevated nitrogen deposition simulation. Following \cite{wieder2019}, who conducted global runs with the default parameterization of the model, we increased CO$_{2}$ to 550ppm, starting in 1998 and for Nitrogen, we added 5gN $m^{-2}$/year, also starting in 1998.

\subsection{Perturbed Physics Ensemble (PPE)}
The CLM5 has a large array of parameters that might have direct and indirect impacts on biogeochemical cycling in the model.  Here we focus on an array of thirteen parameters, chosen for their direct mechanistic impacts on responses to CO$_{2}$ and N fertilization. These parameters are narrowed down from the much larger complete set of CLM5 parameters (which will not enumerate here, since calculation of this number in a model as complex as CLM5 is a scientific exercise in its own right) using a process informed primarily by the model structure, as well as the extensive iterative investigations of the CLM5 model during the development process.  We focus here on biogeochemically important inputs. Parameters relevant to the hydrology and surface energy balance processes (with the exception of the stomatal slope) since sensitivity to these will be assessed by studies currently underway, and parameters of the new plant hydraulics scheme are discussed in \cite{kennedy2019}.

To reduce the complexity involved in understanding and attributing changes in model behaviour with parameter variation, we conducted one-at-a-time (OAAT) perturbations of the parameters around the default state, which (compared to a global parameter sensitivity analysis) allows more straightforward visualization and interpretation of the results.  At each site, we conduct simulations with five values of each parameter (one of which is the default) across the ranges described below. We recognize that a global parameter sensitivity analysis would also be of interest, and indeed have conducted many such analyses. We concluded that OAAT analysis provides clearer insight into the functionality of the parameters with respect to biogeochemical forcing.

\subsection{Default parameter determination}
During the development of CLM5, we conducted extensive testing into the possibility of inverse-calibration of the model using neural network derived emulators, in conjunction with a variety of gridded global data products. We did not use the results of this effort for a variety of reasons, including,  1) the high dimensionality of the space, 2) the degree of non-orthogonal behaviour in the parameter response surfaces, 3) the tendency of the modeled ecosystem to die under lower pre-industrial CO$_{2}$ conditions when calibrated using present-day CO$_{2}$, 4) the dominance of high productivity areas over the calibration of a particular PFT at the expense of marginal areas, and the resulting effect that those marginal areas die off in model spin-up, 5) the subjectivity in weighting of the various observational data products and 6) philosophical issues concerning whether the structural validity of the calibrated model could be independently assessed, given the sparsity of validation datasets. The parameterization used in the default version of the model consequentially utilized the mean trait values where possible from trait datasets (\cite{lawrence2018}) in combination with targeted manual tuning of parameters (including allocation parameters and Nitrogen costs) that were less well constrained by data, to avoid excessively unrealistic values for some model simulated states. We recognize that this more subjective calibration method is less than ideal, and ongoing studies will focus on more limited scope calibration attempts to overcome the difficulties described above, as illustrated by \cite{fer2018}.

@@

The parameters included in our sensitivity analysis are listed in Table \ref{table_ranges}. Their impact on model functionality is described below.  The shorthand version of the parameter name is given in italics.

\textbf{Specific Leaf Area}, \emph{`SLA'}, is a parameter common to almost all land surface models, which determines the area of leaves (in m$^{2}$) derived from one gram of leaf biomass. High SLA values correspond to thinner, 'cheaper' leaves which lead to thicker vegetation canopies and greater interception of light. In CLM5 high SLA leaves also have lower N per unit area (given a target mass-based C:N ratio) and thus lower photosynthestic capacity per unit area.

\textbf{Fine root mass per unit leaf mass}, \emph{`froot\_leaf'}, is the (constant) ratio of fine root biomass to leaf biomass. In CLM5, fine root biomass affects the capacity of vegetation to acquire both water and nutrients via the resistance of the rhizosphere to water flow, and the cost of N acquisition in the FUN model, respectively. Lower values should in principle incur a cost to vegetation in terms of below-ground resource acquisition.  The carbon allocation in CLM5 uses a fixed-fraction approach, with constant ratios between, leaf, fine roots, stem and coarse root allocation. As such, the carbon allocation scheme is thus capable of generating physiologically sub-optimal leaf area index values, for example, with some of the canopy in negative carbon balance. Newer codes within the CLM framework (e.g. the FATES model, https://github.com/NGEET/fates, \cite{fisher2015}m \cite{fisher2018vegetation}) account for this issue by removing leaves in negative carbon balance.

 \textbf{Stem to leaf mass ratio}, \emph{`stem\_leaf'}, controls the fraction of biomass to stem tissues as a fraction of the leaf allocation. Increases in this parameter both decrease the leaf area index (LAI, m$^{2} m^{-2}$) achievable for a given amount of (C and N) allocation to growth, and increase the equilibrium woody biomass. Woody biomass in CLM5 has no functionality and thus there is no cost to plants associated with low values of this parameter).

\textbf{Nitrogen costs}, \emph{`n\_costs'}, are a set of parameters that determine the environmental cost of Nitrogen uptake from the soils. The FUN model introduces the concept of carbon payments for active N uptake, using a bulk approach to account for the costs to plants of low environmental N availability, including; carbon expended on root exudates, symbiotic relationships with soil fungi, expenditure on fine root growth and/or increased metabolic rates to support active uptake. Therefore, uncertainties associated with the cost functions, in particular, are large. These six parameters were defined and calibrated by \cite{brzostek2014}. In CLM5, we maintain the ratios of the parameters as defined by the previous FUN model calibration, but allow the magnitudes of the parameters to change concomitantly, since the definitions of soil N availability vary between CLM5 and \cite{brzostek2014}. The six \emph{'n\_costs'} parameters include \emph{ekc\_active} and \emph{ekn\_active}, the rate constants that relate the cost of ectomycorhhizal uptake to the fine root carbon pool (\emph{ekc\_active}) and soil N concentration (\emph{ekn\_active}), respectively and their counterparts \emph{akc\_active} and \emph{akn\_active} for arbuscular mycorrhizal uptake, and \emph{kc\_nonmyc} and \emph{kn\_nonmyc} for non-mycorrhizal active uptake.

@@

\section{Results}
Our analysis used four sites, with thirteen parameters perturbed to four different values, each of those with control, CO$_{2}$ and N fertilization simulations. Including the default simulations, the total comprises 636 different run configurations. We examine the parametric responses of five output variables chosen for their importance in the overall biogeochemical system (GPP, NPP, LAI, leaf nitrogen  ($N_{area}$, gN m$^{-2}$), and total vegetation carbon (g m$^{-2}$). We did not look at soil output variables as our parameter analysis is focused on vegetation responses.

@@

\subsubsection{Water and N limitation across sites}
The four sites varied in the degree to which they are limited by water and N availability in the control state, with important impacts upon the parametric response characteristics (Figure \ref{btran state}). Water limitation in CLM5 can be expressed via the $B_{tran,mn}$ parameter, which is a model property calculated as a function of the prognostic leaf water potential. $B_{tran}$  is used to reduce assimilation and thus stomatal conductance by scaling the photosynthetic capacity, $V_{c,max}$. Given that leaf water potential and  $B_{tran}$ vary diurnally, we use the average daily minimum value here as a metric of drought stress (1 is no drought stress, 0 is total stomatal closure). For N limitation, the fraction of total NPP (where NPP is GPP-$r_{auto}$) that is used to pay for nitrogen uptake is a direct indicator of how N stress affects the carbon budget.

@@

\subsubsection{Parametric control over system state}
At all sites, specific leaf area (\emph{slatop}) has a large first-order impact on LAI, and on leaf N concentrations (Figure \ref{NR1 state}), as expected since it is used directly in the calculation of both of these quantities. For GPP, the impact of SLA is always positive, despite the higher LAI causing lower N per unit area of leaf (which reduces per-unit leaf area photosynthetic capacity). The impact of SLA on NPP can, however, be negative (e.g., Metolius), since the fixed-fraction allometry model in CLM5 can generate shaded leaves that are in negative carbon balance at high LAI. At Niwot Ridge, the parameter range straddles an optimal SLA (or local maximum at the second-largest parameter values) for NPP, again reflecting the assimilation/respiration cost trade-off.

@@

Changing \emph{leafcn} -the target leaf C:N ratio- over one standard deviation from the mean TRY values, had substantial impacts on all the output variables at all sites. Higher \emph{leafcn} values decrease $N_{area}$, decreasing photosynthetic capacity and hence GPP.  In particular, the higher \emph{leafcn} values had $N_{leaf}$ and photosynthetic capacities so low that GPP was almost zero at Metolius and Harvard Forest and more than halved at Niwot Ridge (\ref{NR1 state}).  We do not find much evidence of saturation in the GPP vs. \emph{leafcn} relationships, in that GPP response does not level off at high leaf N content. This is consistent with the analysis by \cite{lawrence2018} who find that V${_c,max}$ numbers in CLM5 might be low compared to observations.  For NPP,  decreases in respiration rates caused by high \emph{leafcn}, and reduced carbon expenditure on N uptake (Figure \ref{btran state}) somewhat ameliorated the GPP increases, particularly at Caxiuan\~a, which has high ambient temperatures and respiration rates as a result. At Niwot, an optimum \emph{leafcn} ratio (or a local maximum in the parameteric response) was apparent. For NPP this was around the default value, but for LAI it was 0.5 standard deviations lower, reflecting the complex interactions between the expenditure on N uptake and acquisition of N for tissue growth.

@@

Changes in \emph{medlyn\_slope}, the stomatal slope parameter, most directly impact GPP, but can also affect leaf temperature via changes in canopy resistance and thus sensible heat flux. Water limitation is also modified as a result of altered transpiration. At Caxiuan\~a (Figure \ref{CAX state}), the very lowest values of  \emph{medlyn\_slope} caused a complete death of the ecosystem. In Figure \ref{btran state} we can see that low values of the Medlyn stomatal slope ameliorated plant drought stress (by reducing stomatally-mediated water losses), presumably at the cost of limiting photosynthesis past the point at which positive plant growth could be maintained. At Niwot, a clear local maximum was indicated in the response to the Medlyn parameter, indicating that above this level, stomata lose too much water and below it, they do not photosynthesize enough. At Metolius and Harvard forest, only the lower \emph{medlyn\_slope} values affected the photosynthetic rates, indicating co-limitation of GPP by processes other than stomatal conductance.

@@

\subsubsection{Differences across sites}
The responses to CO$_{2}$ and Nitrogen fertilization were substantially different between sites and across parameter ranges.

@@

\subsubsection{Impact of parameters}
Across the four sites, parametric impacts on fertilization responses were very varied, when compared to the relatively consistent impact of parameters on the initial system state. A majority of parameters show no consistent relationship with CO$_{2}$ or N deposition responses across sites, and the range of their impacts varied over an order of magnitude. Thus, the parameter response space of the model is complex with regard to these changing initial conditions, environments, and drivers on global change. Again, we go through the parameters one at a time to highlight the primary features of the model response and how they relate to expectations based on model structure.
